Pillar 1 - Sequential Pattern Mining

### 2019-02-06 Pillar 1 - Sequential Pattern Mining.Rmd
simplify code

# Number of students
n_students <- d_transactions %>%
  select(sequenceID) %>%
  n_distinct

# Probability of taking a course
d_support_take <- d_transactions %>%
  distinct(
    item,
    sequenceID
    ) %>%
  count(
    item
    ) %>%
  mutate(
    rate.take = n / n_students,
    rate.not  = 1 - rate.take
    )

# Probability of failing, having a low grade, having a grade less than x per course
d_support <- d_transactions %>%
  group_by(
    item
    ) %>%
  summarise(
    rate.fail = mean(PF == "fail"),
    rate.low  = mean(HL == "low"),
    rate.0    = mean(Grade <= 0),
    rate.1    = mean(Grade <= 1),
    rate.2    = mean(Grade <= 2),
    rate.3    = mean(Grade <= 3),
    rate.4    = mean(Grade <= 4),
    rate.5    = mean(Grade <= 5),
    rate.6    = mean(Grade <= 6),
    rate.7    = mean(Grade <= 7),
    rate.8    = mean(Grade <= 8),
    rate.9    = mean(Grade <= 9),
    rate.10   = mean(Grade <= 10)
    )

@@

#
# For convenience
make_transaction <- function(data = d_transactions, item = item){

@@

#
# Checking transactions
#inspect(head(transactions_Grade, 10))

@@

#
# For convenience
make_sequence <- function(data = d_transactions, item = item){

  sequences <- data %>%
    filter(!is.na(eventID))%>% #this is here casue we have courses that were never taken.
    group_by(
      sequenceID,
      eventID
    ) %>%
    summarise(
      list_item = list(!!enquo(item))
    ) %>%
    arrange(
      sequenceID,
      eventID
    ) %>%
    ungroup %>%
    pull(
      list_item
      ) %>%
    as("transactions")

  sequences@itemsetInfo <- select(
    data,
    sequenceID,
    eventID
    ) %>%
    as.data.frame

@@

      # Chosen parameters
      parameter = list(
        supp = 0,    # min support
        smax = 1,    # max support
        conf = 0,    # min confidence
        minlen = 2,  # min length of 
        maxlen = 2  # max length of rule
        ),

@@

  type_rule_enquo <- enquo(type_rule)

@@

    # rhs.support
    mutate(
      rhs_course = str_remove_all(rhs, pattern = "_fail|_pass|_low|_not taken")
      ) %>%
    left_join(
      select(data_support, item, !!type_rule_enquo),
      by = c("rhs_course" = "item")
      ) %>%
    mutate(
      rhs.support = !!type_rule_enquo
      )

@@

We encapsulate the three functions `my_apriori()`, `clean_AR()` and `compute_support` into the function `make_AR()`.

@@

```{r AR, cache = TRUE}

#
# AR taken (confidence and lift already correct)
AR_taken <- transactions %>%
  make_AR(
    data_support = d_support_take, 
    type_rule = rate.take
    )

#
# AR fail
AR_PF <- transactions_PF %>%
  make_AR(
    data_support = d_support,
    type_rule = rate.fail
    ) %>%

  # Confidence and Lift
  separate(
    rhs,
    sep = "_",
    into = c("rhs_course", "rhs_outcome"),
    remove = FALSE
    ) %>%
  group_by(
    lhs, rhs_course
    ) %>%
  mutate(
    lhs.rhsTake.support = sum(support),
    confidence          = support / lhs.rhsTake.support,
    lift                = confidence / rhs.support
    ) %>%
  ungroup %>%

@@

#
# AR low
AR_HL <- transactions_HL %>%
  make_AR(
    data_support = d_support,
    type_rule = rate.low
    ) %>%

  # Confidence and Lift
  separate(
    rhs,
    sep = "_",
    into = c("rhs_course", "rhs_outcome"),
    remove = FALSE
    ) %>%
  group_by(
    lhs, rhs_course
    ) %>%
  mutate(
    lhs.rhsTake.support = sum(support),
    confidence          = support / lhs.rhsTake.support,
    lift                = confidence / rhs.support
    ) %>%
  ungroup %>%

@@

#
# AR not taken / fail / pass

AR_TPF <- transactions_TPF %>%
  make_AR(
    data_support = d_support, 
    type_rule = rate.fail
  ) %>%

  # Confidence and Lift
  separate(
    rhs,
    sep = "_",
    into = c("rhs_course", "rhs_outcome"),
    remove = FALSE
    ) %>%
  mutate(
    rhs.take = str_detect(rhs, "fail|pass")
    ) %>%
  group_by(
    lhs, rhs_course
    ) %>%
  mutate(
    lhs.rhsTake.support = sum(support * rhs.take),
    confidence          = support / lhs.rhsTake.support,
    lift                = confidence / (rhs.support + 1e-4) # increase denominator by smal value to avoid dividing by 0.
    ) %>%
  ungroup %>%

@@

rm(transactions, transactions_PF, transactions_HL, transactions_TPF)
```

>>>>

#
# Number of students
n_students <- d_transactions %>%
  select(sequenceID) %>%
  n_distinct

#
# Probability of taking a course
d_support_take <- d_transactions %>%
  distinct(
    item,
    sequenceID
    ) %>%
  count(
    item
    ) %>%
  mutate(
    rate.take = n / n_students,
    rate.not  = 1 - rate.take
    )

#
# Probability of failing, having a low grade, having a grade less than x per course
d_support <- d_transactions %>%
  group_by(
    item
    ) %>%
  summarise(
    rate.fail = mean(PF == "fail"),
    rate.low  = mean(HL == "low"),
    rate.0    = mean(Grade <= 0),
    rate.1    = mean(Grade <= 1),
    rate.2    = mean(Grade <= 2),
    rate.3    = mean(Grade <= 3),
    rate.4    = mean(Grade <= 4),
    rate.5    = mean(Grade <= 5),
    rate.6    = mean(Grade <= 6),
    rate.7    = mean(Grade <= 7),
    rate.8    = mean(Grade <= 8),
    rate.9    = mean(Grade <= 9),
    rate.10   = mean(Grade <= 10)
    )

@@

make_transaction <- function(data = d_transactions, item = item){

@@

#
# Checking transactions
# inspect(head(transactions, 10))

@@

make_sequence <- function(data = d_transactions, item = item){

  sequences <- data %>%
    filter(
      !is.na(eventID) #this is here casue we have courses that were never taken.
      ) %>% 
    group_by(
      sequenceID,
      eventID
    ) %>%
    summarise(
      list_item = list(!!enquo(item))
    ) %>%
    arrange(
      sequenceID,
      eventID
    ) %>%
    ungroup %>%
    pull(
      list_item
      ) %>%
    as("transactions")

  # indicating sequence ID and event ID for each transaction
  sequences@itemsetInfo <- select(
    data,
    sequenceID,
    eventID
    ) %>%
    as.data.frame

@@

      # include all AR possible
      parameter = list(
        supp   = 0, # min support
        smax   = 1, # max support
        conf   = 0, # min confidence
        minlen = 2, # min length of rule
        maxlen = 2  # max length of rule
        ),

@@

@@

    # rhs.support
    mutate(
      rhs_course = str_remove_all(rhs, pattern = "_fail|_pass|_low|_not taken")
      ) %>%
    left_join(
      select(data_support, item, !!enquo(type_rule)),
      by = c("rhs_course" = "item")
      ) %>%
    mutate(
      rhs.support = !!enquo(type_rule)
      )

@@

We encapsulate the four functions `my_apriori()`, `clean_AR()`, `compute_support()` and `editing_AR()` into the function `make_AR()`.

@@

```{r AR take}
AR_taken <- transactions %>%
  make_AR(
    data_support = d_support_take, 
    type_rule = rate.take
    )

rm(transactions)
```

```{r AR fail}
AR_PF <- transactions_PF %>%

  make_AR(
    data_support = d_support,
    type_rule = rate.fail
    ) %>%

  # Confidence and Lift by hand
  separate(
    rhs,
    sep = "_",
    into = c("rhs_course", "rhs_outcome"),
    remove = FALSE
    ) %>%
  group_by(
    lhs, rhs_course
    ) %>%
  mutate(
    lhs.rhsTake.support = sum(support),
    confidence          = support / lhs.rhsTake.support,
    lift                = confidence / rhs.support
    ) %>%
  ungroup %>%

@@

rm(transactions_PF)
```

```{r AR low}
AR_HL <- transactions_HL %>%

  make_AR(
    data_support = d_support,
    type_rule = rate.low
    ) %>%

  # Confidence and Lift by hand
  separate(
    rhs,
    sep = "_",
    into = c("rhs_course", "rhs_outcome"),
    remove = FALSE
    ) %>%
  group_by(
    lhs, rhs_course
    ) %>%
  mutate(
    lhs.rhsTake.support = sum(support),
    confidence          = support / lhs.rhsTake.support,
    lift                = confidence / rhs.support
    ) %>%
  ungroup %>%

@@

rm(transactions_HL)
```

```{r AR TPF}
AR_TPF <- transactions_TPF %>%
  make_AR(
    data_support = d_support, 
    type_rule = rate.fail
  ) %>%

  # Confidence and Lift
  separate(
    rhs,
    sep = "_",
    into = c("rhs_course", "rhs_outcome"),
    remove = FALSE
    ) %>%
  mutate(
    rhs.take = str_detect(rhs, "fail|pass")
    ) %>%
  group_by(
    lhs, rhs_course
    ) %>%
  mutate(
    lhs.rhsTake.support = sum(support * rhs.take),
    confidence          = support / lhs.rhsTake.support,
    lift                = confidence / rhs.support
    ) %>%
  ungroup %>%

@@

rm(transactions_TPF)
```
